Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2020-12-10 GWLDCI.tex
and a spellcheck

In \cite{wilsonloop} Agarwala and Marin-Amat showed that every weakly admissible Wilson loop diagram can be identified with a type of matroid called a positroid, and observed that this mapping is neither one-to-one nor onto. Each such positroid corresponds to a positroid cell in the appropriate positive Grassmannian $\Gr(k,n)$ \cite{Postnikov}. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the faces up to paralellism in an associahedron.

@@

\begin{proof}
  Since $V_{>3}(D_1) = V_{>3}(D_2)$, the vertices of the non-triangle faces of $D_1$ and $D_2$ are the same, but $D_1$ and $D_2$ do not have the same set of non-triangle faces.
  Suppose that no diagonal of $D_1$ crosses through a non-triangle face of $D_2$.  Then all non-triangle faces of $D_2$ must lie strictly within faces of $D_1$.  Since the non-trangles faces of the two dissections are different at least one of these containments must be proper, and so some diagonal bounding one of the non-triangle faces of $D_2$ must cross through a non-triangle face of $D_2$.  Therefore, swapping $D_1$ and $D_2$ if necessary,  
  %Since $D_1$ and $D_2$ do not have the same set of non-triangle faces, but the vertices of the non-triangles faces are the same, $V_{>3}(D_1) = V_{>3}(D_2)$,
  there must be some diagonal $d$ of $D_1$ that crosses through a non-triangle face $f$ of $D_2$.
  If $f$ has degree at least 5 then it must contain four vertices in the configuration described in the statement.

@@

In summary, using matroids and polygon dissections we have counted Wilson Loop diagrams that are equivalent in the sense of having the same positroid, and characterized Wilson Loop diagrams that are inequivalent in terms of faces of the associahedron $A_n$ up to paralellism.  The second paper in this series calculates the dimension of the positroid cell associated to a Wilson loop diagram in a very concrete way by counting plusses in a Le diagram, and looks at the denominator of the integrand of the Wilson Loop diagram.  All of these results are interesting mathematics while also serving to better understand the physics of scattering amplitudes.

>>>>

In \cite{wilsonloop} Agarwala and Marin-Amat showed that every weakly admissible Wilson loop diagram can be identified with a type of matroid called a positroid, and observed that this mapping is neither one-to-one nor onto. Each such positroid corresponds to a positroid cell in the appropriate positive Grassmannian $\Gr(k,n)$ \cite{Postnikov}. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the faces up to parallelism in an associahedron.

@@

\begin{proof}
  Since $V_{>3}(D_1) = V_{>3}(D_2)$, the vertices of the non-triangle faces of $D_1$ and $D_2$ are the same, but $D_1$ and $D_2$ do not have the same set of non-triangle faces.
  Suppose that no diagonal of $D_1$ crosses through a non-triangle face of $D_2$.  Then all non-triangle faces of $D_2$ must lie strictly within faces of $D_1$.  Since the non-triangle faces of the two dissections are different at least one of these containments must be proper, and so some diagonal bounding one of the non-triangle faces of $D_2$ must cross through a non-triangle face of $D_2$.  Therefore, swapping $D_1$ and $D_2$ if necessary,  
  %Since $D_1$ and $D_2$ do not have the same set of non-triangle faces, but the vertices of the non-triangles faces are the same, $V_{>3}(D_1) = V_{>3}(D_2)$,
  there must be some diagonal $d$ of $D_1$ that crosses through a non-triangle face $f$ of $D_2$.
  If $f$ has degree at least 5 then it must contain four vertices in the configuration described in the statement.

@@

In summary, using matroids and polygon dissections we have counted Wilson Loop diagrams that are equivalent in the sense of having the same positroid, and characterized Wilson Loop diagrams that are inequivalent in terms of faces of the associahedron $A_n$ up to parallelism.  The second paper in this series calculates the dimension of the positroid cell associated to a Wilson loop diagram in a very concrete way by counting pluses in a Le diagram, and looks at the denominator of the integrand of the Wilson Loop diagram.  All of these results are interesting mathematics while also serving to better understand the physics of scattering amplitudes.
